---
title: "Notes - Things They Forgot to Teach You In R, rstudio::conf18"
author: "Jessica Minnier"
date: "1/31/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thanks to the excellent contributions (via `pull` request!) from classmate [Peter Higgins!](https://github.com/higgi13425)

# Links:

Materials: [rstd.io/forgot](https://github.com/jennybc/what-they-forgot)

To use "in building" mirror: `options(repos = c(CRAN = "https://cran.rstudio.com/"))`

# Random thoughts

## Day 1: Morning - library exploration

-   JB has us go to issues and put emoji on OS, tests that we've signed into github! also sees how many windows users there are
-   it's ok that things take a long time
-   make things self explaining (good filenames - see https://speakerdeck.com/jennybc/how-to-name-files), don't spend a lot of time writing wordy explainers that you have to maintain and won't want to read later
-   organize your files and keep your readme up to date!
-   "unless you can keep it current (maintain it), don't write it"
-   it helps to know about your R installation and where packages go in order to make a package
-   new package `fs` (for 'file system'), helps us work with file paths. note, don't use paste() to work with file paths like they are strings, they are not!

## Day 1: Mid-morning - file copying and naming, projects

-   turn off .RData loading and never ask to save .RData
-   restart R (Shift-Cmd-F10) often!
-   no absolute paths
-   use a stable base: `here::here("data","raw-data.csv")` in your project directory, or `fs::path_home()`
-   form paths at runtime relative to a stable base
-   normalizePath() makes it work on both windows and mac
-   Rstudio set up a mirror for downloading packages **in** the building! Whaaattt!
-   use `basename` instead of `strsplit` to get name of file from a path (don't have to specify  or /)
-   name files and directories as thing_thing-info_thing2 so for instance day_session_topic so you can order and delimit them later

## Side note from my own issues in moving this repo to github

How to add an existing Rstudio repo to github:

This has changed since [git 2.9](https://stackoverflow.com/questions/37937984/git-refusing-to-merge-unrelated-histories), now when you add remote and then pull you need a special option. This only needs to happen if you added a Readme. Just don't do this and life will be easier.

```
git remote add origin https://github.com/jminnier/rstudioconf18_jennybryan_forgot.git
git pull origin master --allow-unrelated-histories
```

Most easy option from [happywithgitr](http://happygitwithr.com/existing-github-last.html)

```{r, eval=FALSE}
usethis::use_github()
```

## Day 1: Afternoon - git/github

-   git is scary but you should learn the 3 things you always have to do (commit, push, pull) and just do them over and over until you are comfortable ("get off the beach!")
-   easier to start from github then go to rstudio (sad)
- if you are more of a visual thinker, you might like useing a graphical front end for git, like SourceTree or GitKraken
-   when you clone a repo in Rstudio and copy the ssh (or https) into the prompt, then press TAB, it will auto-fill the name of the repo into the next field!
- if you want to collaborate on someone else's repo, best practice to (1) fork their repo so you have a copy in your GitHub (2) then clone this to your local git repo. This will keep links, so that later you can push your changes to your GitHub repo and then issue a pull request to original owner.
-   in git commit: two yellow question marks: local file git has never seen before
-   jenny commits her .Rproj (this is controversial)
-   intermediate step "staging" tells git you do want these files to be part of the next commit, this way you can commit files separately. 
- Staging - put things that are ready to mail in a box, Commit - mail the whole box.
-   use conventional file extensions so that github can show you customized diffs
-   jenny thinks putting derived products in version control is a good idea (i.e. html from Rmds), but not binary files (word files, pdfs)
-   jim hester will show us how github makes it easy to search code
-   when using workbook button from .R files we get a html report (jenny says html maybe not best option, more later)
- note to knit to pdf, start by installing tinytex package, then run tinytex::install_tinytex, then you are able to knit to PDF
-   oh yeah, html does not look good on github =( too raw!
-   mullet: what you need to write (.R, .Rmd) = this business in front; what you want to read/see (.html, .md) = party in the back (or is it the other way around?)
-   .html is not useful but .md is very useful for github; use YAML:
```
#' ---
#' output: github_document
#' ---
```
-   now github will render the resulting .md file this to look pretty nice!
-   source: only hard-liners say don't version control output (only source code). but having .md files or .html files makes it really useful to see what the output looks like and how it changes.
-   JB takeaway 1: if you want to make your stuff immediately consumable, rendering to markdown and commiting that at the same time as the R file, gives something others can look at with little to no effort (be kind and be realistic!)
-   JB takeaway 2: markdown is vastly more useful than .html and .pdf (on github)
-   More about github_documents [here](http://rmarkdown.rstudio.com/github_document_format.html)
-   files that look nice in github: .csv, .png, .md, see the browsability section of happywithgitr


## Day 1: Afternoon after break

-   to upload an existing git directory to github without going through all the steps, just do `usethis::use_github()`!!!
-   JB showing that .R and .Rmd can make same github_document output
- You can make your README.md into a webpage by changing settings in GitHub. Go to your repo, click 'Settings', scroll way down to GitHub pages, change Source to Master, and choose a theme. Can add links to other .md files in the same repo by inserting this into README.md file: [displayname](link.md)
-   Good options for knitr:
```{r, eval=FALSE}
knitr::opts_chunk$set(
          collapse = TRUE,
          comment = "#>",
          out.width = "100%"
)
```
-   git can detect simple push conflicts (i.e. you forgot to pull first)
-   if you are collaborating on code: commit and push (and pull) often, so that you make small changes that git's troubleshooting push fail algorithm can handle it; "sync to the mothership really often"
-   also working with binary files like pdf makes this really hard since git can't look inside them to figure out what the problem is. so, don't put them in your repo very often! (.gitignore)
-   use githooks to avoid pushing files that have conflicts (they can look for the markers from the merge conflict)
- if you have made conflicting commits to your local version and remote version, push will be rejected. Pull first and fix conflicts. If this does not work, consider git reset. If this does not work, consider renaming local repo to borked-repo, pull remote, then selectively copy over new useful files from borked-repo
-   you can search github for code snippets (I need to do this more!), such as: ["llply" user:cran language:R](https://github.com/search?l=r&q=%22llply%22+user%3Acran+language%3AR&ref=searchresults&type=Code&utf8=%E2%9C%93); Jim Hester will talk more about this tomorrow


## Day 2 Morning - Jim Hester, searching github

- Jim Hester is teaching us how to use github search, for example `read_csv OR read_delim col_type language:R extension:.R` looks for examples of using the `col_type` argument in `read_csv` or `read_delim` functions
- if the documentation doesn't answer your questions, use github search to find examples
- example search: `user:cran onAttach` will search CRAN packages that use the function `onAttach`
- use `user:` field to search CRAN or tidyverse because those users/orgs keep track of all the packages on cran or tidyverse
- if you get an error, you can search that error code string in the repo to determine what is causing it
- The tidyverse packages typically name internal helpers with a trailing underscore.
- JH has a package for this! [jimhester/lookup](github.com/jimhester/lookup)
```{r, eval=FALSE}
library(lookup)
lookup(knitr::knit)
```
- JB: if you are reporting a bug or requesting an addition, be as specific as possible

## Day 2 morning, JB start up files

- re: installing packages after an update: you can install them organically as you go, this is a good way to perform "spring cleaning" (JB and [@LucyStats](twitter.com/lucystats) are in favor of this). However, Jim Hester has a package for this, too! [jimhester/autoinst](github.com/jimhester/autoinst) (but maybe try not to break it/find issues ;))
- github search for "dotfiles" shows people's repos of dotfiles
- very useful links from JB's repo:
    + Kevin Ushey blog post on `.Rprofile` [RProfile Essentials](http://kevinushey.github.io/blog/2015/02/02/rprofile-essentials/)
    + Kevin's [etc repo](https://github.com/kevinushey/etc) for dotfiles and the like
    + Shaun Jackman's [dotfiles repo](https://github.com/sjackman/dotfiles)
- examples of good options to turn on:
    + `warnPartialMatchAttr = TRUE`, etc = no autocomplete!
    + `utils::rc.settings(ipck=TRUE)`  = yes autocomplete package names
- don't load packages in .Rprofile because you will be using functions other people can't without loading
    + though JB still loads "workflow packages" since low risk for irreproducible code: devtools, usethis, remotes, testthat, reprex
- can populate description file automatically (see JB's .Rprofile)
- interesting package to maintain these types of files: [HenrikBengtsson/startup](https://github.com/HenrikBengtsson/startup)

## Day 2, Morning after break

- github etiquette: use `#` in front of numbers to link to issues, words like "Fixes" are keywords that can close issues [see more on github](https://help.github.com/articles/closing-issues-using-keywords/)
- plan for revisiting analyses/projects (like 10x more than you thought you would)
- refine and extend! make your code more readable, efficient, resilient, general
- beware of monoliths! i.e one script and a workspace, much better to break logic and ouput into smaller pieces; divide and conquer; i.e. making figures is separate from modeling
- make scripts based on the different steps in the analysis pipeline (see slide 17 [slides/day2_slides2_project-api.pdf](slides/day2_slides2_project-api.pdf))
- make outputs separate based on their uses rather than a giant RData file
- practice "safe paths"! use `here` package to do this; works on my machine and on yours, also works with knitr and rmarkdown! rmarkdown sets workign directory where .Rmd file is, but `here` can get around this if you have subdirectories
```{r, eval=FALSE}
library(here)
here()
here("day1_s1_explore-libraries")
cat(readLines(here("day1_s2_copy-files/","00_filesystem-practice_comfy.R")))
```
- we had an excerise of forking a repo and adding code to files along with documentation in the readme.
- there are R packages that are similar to `make`, such as remake (more to come after lunch)


# Q's

-   what about versioning data? for small to medium you can easily use github, might need to just keep it in a separate repo and git that, or might need to use gitfs for large file systems
-   in the analysis pipeline, if I want to generate an html (or pdf etc) file along with the github_document, is there a way to always generate both? you can "keep md" but can you "keep github_document"? Sort of, but you need to use `rmarkdown::render` (in theory, though JB can't get it to work yet for github_docs...)
-   why is `fs` package making my Rstudio crash every time I want to view an object created with it? May be related to this [issue](https://github.com/r-lib/fs/issues/58)

